### 2020-08-21 coverletter.tex

\documentclass{letter}

\signature{Diego Muñoz-Gacitúa}

\begin{document}

\begin{letter}{Diego Muñoz-Gacitúa\\ Facultad de Ciencias \\ Universidad de
    Chile \\ Las Palmeras \#3425, Ñuñoa, Chile}

\opening{Dear referee,}
I am pleased to submit an original research article entitled ``Characterization
and validation of a membrane mimetic with magnetic orienting capabilities and
high phospholipid content'' to be considered in publication in the Journal of
Molecular Liquids.\\

In this manuscript we develop a membrane mimetic which imitates the permeability
of a natural cellular membrane, this property was tested by confirming that
Benzocaine, a local anesthetic that is proven to permeate the cellular membrane,
does permeate the membrane mimetic proposed; and by confirming that Levodopa, an
amino-acid that is unable to permeate the cellular membrane via passive
mechanisms, cannot permeate the membrane mimetic. We also provide a molecular
dynamics simulation model calibrated to reproduce experimental $^2$H-NMR results
on the mimetic.\\

We believe this manuscript is appropriate for publication in the Journal of
Molecular Liquids because our membrane mimetic can be of interest to the
readers of the journal seeking advances in lyotropic liquid crystals and biomolecules in solution.\\

This manuscript is an original work that has never been published
before, and is not under consideration by any other journal. We have
no conflicts of interest to disclose.

\closing{Sincerely,}

\end{letter}
\end{document}