### 2020-08-21 revision_cover_letter.tex

\documentclass{letter}

\signature{Diego Muñoz-Gacitúa}

\begin{document}

\begin{letter}{Diego Muñoz-Gacitúa\\ Facultad de ciencias \\ Universidad de
    Chile \\ Las Palmeras \#3425, Ñuñoa, Chile}

  \opening{Dear Prof. Schroer}

   We are greateful for your consideration of our manuscript titled
   ``Characterization and validation of a membrane mimetic with magnetic
   orienting capabilities and high phospholipid content'' to be published in The
   Journal of Molecular Liquids as a full article.

   We would also like to extend our thanks to both of the reviewers, who have
   taken their time to give such a deep and extensive review of our work.

   \closing{Sincerely,}

 \end{letter}
 \end{document}